% Committee
First and foremost, I would like to thank my advisor, Ed Suh. Without Ed's
guidance, collaboration, and mentorship, this thesis would not have been
possible. I would also like to thank my other committee members, Rajit Manohar and Mark
Campbell, for their guidance and advice throughout my Ph.D. career.

% Collaborators
The work in this thesis was not a solo effort. Special thanks to Tao Chen and
Mohamed Ismail for their help on our work on secure and reliable real-time
systems. Similarly, special thanks to Taejoon Song for his help on our work on
energy-efficient real-time systems.  I want to thank Rahul Mangharam for our
early discussions that led us to pursue this research direction into real-time
systems.  I would also like to thank Fran\c{c}ois Guimbreti\`{e}re for his
feedback and insight on user interactions. 

% Academic Community
The Suh Research Group and CSL have been a great academic, as well as
social, community to be a part of. They have been a wonderful group of people
to bounce ideas of, share knowledge, and to get lunch with. 
Thank you to Ray Huang for helping me get set up and started when I first
arrived at Cornell.  Thank you to KK Yu for always being there whenever I
needed a favor. Thanks to Ji Kim for introducing to me to Dungeons and Dragons
and a myriad of new board games.  Thanks to Andrew Ferraiuolo for always being
free when I needed to take a break and thanks to Shreesha Srinath for showing me
some great places to eat in Ithaca.

% Friends
My time at Cornell would not have been as enjoyable without all the friends that I have made here.
Thank you to Nata Saslafsy, Koyuki Nakamura, Harrison Ko, and Scott Johnson for
some great skate sessions and for teaching me to snowboard. 
Thanks to Crystal Lu for showing me that there's more to life than work.
Thanks to Carissa Kang, Elizabeth Wijaya, Lynette Saw, Bryan Chong, and Mairin
Ng for teaching me Singlish over Mahjong.
I also want to thank the Cornell Longboarding Club, Libe Slope FC, Mystery
Machine, the Cayuga Windsurfing Club, the Cornell Cube Clube, the Glen Region
SCCA, and the Enchanted Badger. These communities have helped me make Ithaca a
place to call home for these six years.

% CSL/TSG
%The Suh Research Group has had a large contribution to my graduate career. The
%group has provided me with feedback and discussion on my work. I would like to
%specifically thank Tao Chen and Mohamed Ismail for their help with our work on
%design secure and reliable real-time systems. I would also like to thank
%Taejoon Song for his help on our work on energy-efficient real-time systems.
%I want to thank CSL as a whole for being a great group to bounce ideas off and
%to learn about other areas of related research. They have served as a great
%academic community. Thanks to Ji Kim, KK Yu, Andrew Ferraioulo, Jon Tse, and
%others for some great board game, Dungeons and Dragons, and Netrunner sessions.

% Family
Finally, I would like to thank my parents for teaching me the value of hard
work and discipline. I would not have been able to complete this journey
without their continued support.

